---
title: "Interactive Section"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: Exercises

---

<style>
  div.yellowbox {
    background-color: #f1dfa6;
    border: 1px solid black;
    padding: 21px;
  }

  strong {
    color: #d10056
  }

  .center {
    text-align: center;
  }
</style>


```{r setup, include=FALSE}
library(learnr)
library(rio)
library(tidyverse)
library(tidymodels)
library(janitor)
library(kableExtra)

knitr::opts_chunk$set(exercise=TRUE, exercise.eval=FALSE, exercise.lines=6, attr.source = ".numberLines")
```

:::: {.yellowbox data-latex=""}

::: {.center data-latex=""}
**How to Work with this Interactive Section**
:::

Here you find the complete interactive section from the book, including book content, and R-code-blocks. Read the content in a browser as you would read it in the textbook. 
With the R-code blocks, you can work in various ways:

* Use the `Run Code` button to execute the R-code and see the result.

* Change the R-code, click the `Run Code` button, and then see the new results.

* To get the original R-code back, click the `Start Over` button at the left-top corner of each R-code block. 

* To reset the R-code for all code blocks, click the `Start Over` in the left menu.

**Make sure you installed all needed packages!** See Section "R Packages Required"  in this Chapter for a list of packages that need to be installed. Section 3.2 explain how to install a packege in *RStudio*.



**For Troubleshooting:** [Click here](https://blog.lange-analytics.com/2024/01/interactsessions.html)
::::


 
 
 <clex id="Exc100">

### Introduction: Was Chivalry Dead on the Titanic? \index{Chivalry Dead Exercise}

In this interactive section, you will work on a small project to determine if chivalry was dead when the *Titanic* sank in 1912[^03-randrstudio-17]. The background for the project is an article written by Hannah Furness (Royal Correspondent at The Telegraph).[^03-randrstudio-18] She cites a Swedish historian claiming, "It was 'not true' that women and children on board the *Titanic* survived thanks to the gallantry of men."[^03-randrstudio-19] Hannah Furness also reports anecdotal evidence about men being in the lifeboats instead of women: "Over on the starboard side, it was a different story. On the last boats to leave, the majority were men."[^03-randrstudio-20] However, Furness did not use any statistics to support her claim, although data are publicly available.

[^03-randrstudio-17]: The website History.com (keyword: *Titanic*) provides a comprehensive article about the *Titanic* disaster (see @History2021)

[^03-randrstudio-18]: See @Furness2022.

[^03-randrstudio-19]: See @Furness2022.

[^03-randrstudio-20]: See @Furness2022.

It is now your job to fill this gap. The question is:

**Was the proportion of women surviving the Titanic disaster greater than the one of men?**

If true, that would strongly indicate that the ultimate chivalry --- dying to protect a woman's life --- was still present when the *Titanic* sank and that Furness' claim that chivalry was dead in 1912 is likely incorrect.

### Analyzing Titanic Survival Rates for Women and Men

```{r RAndRStudio-Exerc100-DataChivalryLoad, exercise=FALSE, include=FALSE}
library(tidyverse);library(rio)

DataTitanic=import("https://jomanab-cloud.github.io/ISE351-25-Presentations/Titanic.csv") 

DataAnalysis=DataTitanic |>
             as_tibble() |> 
             select(Survived, Gender, PasClass=Pclass) |>
             mutate(Survived=ifelse(Survived==1,TRUE,FALSE))
DataWomen=DataAnalysis |> 
          filter(Gender=="female")
DataMen=DataAnalysis |> 
          filter(Gender=="male")
FemTotal=nrow(DataWomen)
FemSurv=sum(DataWomen$Survived)
MaleTotal=nrow(DataMen)
MaleSurv=sum(DataMen$Survived)
FemSurvRate=mean(DataWomen$Survived)
MaleSurvRate=mean(DataMen$Survived)

```

You will use the same *Titanic* dataset we used before together with the `tidyverse` package. The *Titanic* dataset contains 887 observations about *Titanic* passengers, including variables such as a survival indicator, passenger's gender, and the class traveled in. The dataset covers about 67% of the total passengers from the *Titanic*.

Below are the first six observations from the `DataTitantic` data frame:

```{r RAndRStudio-Exerc100-100, exercise=FALSE, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse);library(rio)
DataTitanic=import("https://jomanab-cloud.github.io/ISE351-25-Presentations/Titanic.csv") 
head(DataTitanic)
```

```{r, RAndRStudio-Exerc100-LoadDataForNoExerc, exercise=FALSE, eval=TRUE, echo=FALSE}

DataAnalysis=DataTitanic |>
             as_tibble() |> 
             select(Survived, Gender, PasClass=Pclass) |>
             mutate(Survived=ifelse(Survived==1,TRUE,FALSE))
DataWomen=DataAnalysis |> 
          filter(Gender=="female")
DataMen=DataAnalysis |> 
          filter(Gender=="male")
FemTotal=nrow(DataWomen)
FemSurv=sum(DataWomen$Survived)
MaleTotal=nrow(DataMen)
MaleSurv=sum(DataMen$Survived)
FemSurvRate=mean(DataWomen$Survived)
MaleSurvRate=mean(DataMen$Survived)
```

Before writing *R* code, we need to develop a **data analysis strategy**:

**Step 1:**

:   You will create a data frame, `DataAnalysis`, containing only the variables $Survived$, $Gender$, $Pclass$. In addition, you will transform the variable $Survived$ to a logic variable with `TRUE` and `FALSE` values, and you will rename $Pclass$ to $PasClass$.

**Step 2:**

:   You will divide the data frame `DataAnalysis` into two new data frames, `DataWomen` and `DataMen,` containing only observations for women and men, respectively.

**Step 3:**

:   The data frames `DataWomen` and `DataMen` will finally be used to calculate the survival rates of women and men.

Remember, whenever you work on a data science project, develop a detailed data analysis strategy before you start coding. Otherwise, after you have spent a lot of time coding, you might need to entirely change your code because you forgot to consider an important fact.

You will now follow the steps outlined above to write the *R* code for the project. **Step 1** from the data analysis strategy requires to create a new data frame `DataAnalysis` that only contains the variables `Survived` (as `logic` data type), `Gender`, and `Pclass` (renamed to `PasClass`).

When you substitute the `...` in the code below (note that the data frame `DataTitanic` has already been loaded in the background), the command `head(DataAnalysis`) will print the first six observations of the new data frame to the console:

```{r RAndRStudio-Exerc100-200, exercise=TRUE, eval=FALSE, exercise.setup="RAndRStudio-Exerc100-DataChivalryLoad"}
DataAnalysis= DataTitanic |>
               select(... ,... , ... = ...) |>
               mutate(Survived=as.logical(Survived))
head(DataAnalysis)
```

```{r RAndRStudio-Exerc100-200-hint, eval=FALSE, exercise=FALSE, include=FALSE, exercise.lines=11}
'HINT: Use the select() command and list the variables (columns) that you like to
have included in the new data frame. Make sure to rename Pclass to PasClass.
  
The command to convert the variable Survived from numeric to logic 
is already provided with hte mutate command.

The next HINT provides the solution. So, keep trying before you peek.'
```

```{r RAndRStudio-Exerc100-200-solution, eval=FALSE, exercise=FALSE, include=FALSE}
DataAnalysis=DataTitanic |>
               select(Survived,Gender,PasClass=Pclass) |>
               mutate(Survived=as.logical(Survived))
head(DataAnalysis)
```

Note that in the code above, the variable `Survived` was converted from data type `numerical` to `logic` with the `mutate()` command: After `as.logical()` converted `Survival` into `logical` data type, the result was assigned to the variable `Survival` updating it with the new data type.

**Step 2** from the data analysis strategy requires splitting the data frame `DataAnalysis` into two data frames:

1)  `DataWomen` should only contain female passengers.
2)  `DataMen` should only contain male passengers.

In the code block below you can use the `filter()` command to generate the two new data frames (note that the data frame `DataAnalysis` has been loaded already in the background):

```{r RAndRStudio-Exerc100-300, exercise=TRUE, eval=FALSE, exercise.setup="RAndRStudio-Exerc100-DataChivalryLoad"}
DataWomen=DataAnalysis |>
          filter(... == ...)
DataMen=DataAnalysis |>
          filter(... == ...)
head(DataWomen)
head(DataMen)
```

```{r RAndRStudio-Exerc100-300-hint, eval=FALSE, exercise=FALSE, include=FALSE, exercise.lines=11}
'HINT: The two filter() commands need  a condition for
the observations to be filtered.

Add the variable name left of the "==" and the filter-value 
right of the "==" (do not forget the quotes).

The next HINT provides the solution. So, keep trying before you peek.'
```

```{r RAndRStudio-Exerc100-300-solution, eval=FALSE, exercise=FALSE, include=FALSE}
DataWomen=DataAnalysis |>
          filter(Gender=="female")
DataMen=DataAnalysis |>
          filter(Gender=="male")
head(DataWomen)
head(DataMen)
```

Now that you have two different data frames that either only contain male or female passengers' observations, you are ready to calculate the related survival rates.

In the code block below, you will calculate the female survival rate. Later, for the male survival rate, we will show you a shorter alternative.

The female survival rate $(Fem_{SurvRate})$ is defined as:

$$Fem_{SurvRate}=\frac{Fem_{Surv}}{Fem_{Total}}$$ Therefore, to calculate the female survival rate, you need to find the number of female passengers that survived $(Fem_{Surv})$ and the total number of female passengers $Fem_{Total}$. The latter is the same as the number of observations in the data frame `DataWomen` and can be calculated with `nrow(DataWomen)`.

To find the number of surviving women $(Fem_{Surv})$, we can use the fact that a surviving passenger was coded with `TRUE` with an internal value of $1$ and a not-surviving passenger with `FALSE` with an internal value of $0$. Therefore, when we extract the variable $Survived$ as a vector-object from the data frame 'DataWomen' and calculate the sum from this vector, we get the number of female survivors `(sum(DataWomen$Survived))`.

With this information, you should be ready to complete the code below to calculate the rate of female survivors (note that the data frame `DataWomen` has been loaded already in the background):

```{r RAndRStudio-Exerc100-350, eval=FALSE, exercise=TRUE, exercise.setup="RAndRStudio-Exerc100-DataChivalryLoad"}
FemSurv=...(...$...)
FemTotal=...(...)
FemSurvRate= FemSurv/FemTotal
cat("The survival rate for wommen is:",FemSurvRate)
```

```{r RAndRStudio-Exerc100-350-hint, eval=FALSE, exercise=FALSE, exercise.lines=11, include=FALSE}
'The female survival rate is calculated by dividing the number of female
survivors by the total number of female passengers.

The sum of the variable Survived in the data frame DataWomen gives you the
number of female survivors (why?).

The number of female passengers is the same as the number of observations in
the data frame `DataWomen`, which you can get with the command nrow().

The next HINT provides the solution. So, keep trying before you peek.'
```

```{r RAndRStudio-Exerc100-350-solution, eval=FALSE, exercise=FALSE, exercise.setup="RAndRStudio-Exerc100-DataChivalryLoad", include=FALSE}
FemSurv=sum(DataWomen$Survived)
FemTotal=nrow(DataWomen)
FemSurvRate= FemSurv/FemTotal
cat("The survival rate for women is",FemSurvRate)
```

You could do the same for the male passengers. However, there is a shorter alternative to calculate the male survival rate (note that the data frame `DataMen` has been loaded already in the background):

```{r RAndRStudio-Exerc100-400, exercise=FALSE, eval=TRUE}
MaleSurvRate= mean(DataMen$Survived)
cat("The survival rate for men is",MaleSurvRate)
```

Why could the `mean()` command be used to calculate the survival rate directly? It is because the *mean* is calculated by dividing the sum of a variable by the number of observations. This is what we did explicitly when we calculated the female survival rate.

When we use the `cat()` command to print and compare the female and male survival rates, you can see that the survival rate for female passengers is much higher than the one for female passengers:[^03-randrstudio-21]

[^03-randrstudio-21]: The character value "\\n" creates a line break in the console output

```{r RAndRStudio-Exerc100-410, exercise=FALSE, eval=TRUE}
cat("The survival rate for women is:", FemSurvRate, "\n",
    "The survival rate for men is:", MaleSurvRate )
```

The remaining question is if the difference between the two survival rates is high enough to allow us to reject the assertion of the *Telegraph* article that *chivalry was dead in 1912*?

To answer this question, we need more analysis. *R* really shines when data are well organized, as done here --- when good data engineering is performed. With the `prop.test()` command, *R* can run a test if the survival rates for women and men differ significantly. The syntax only requires the values for four arguments:

-   the number of female survivors,

-   the number of male survivors,

-   the number of total female passengers, and

-   the number of total male passengers.

These values have been calculated already and stored in the *R* Objects `WomenSurv`, `MenSurv`, `WomenTotal`, and `MenTotal`. We can provide these values to the `prop.test()` command grouped as two *R* vectors, one for the survival numbers and one for the total numbers for women and men, respectively:

```{r RAndRStudio-Exerc100-700, exercise=FALSE, eval=TRUE}
ModelPropFemaleVsMale=prop.test(c(FemSurv, MaleSurv),
                                c(FemTotal, MaleTotal))
```

Above, we saved the output of the `prop.test()` command in an *R* object named `ModelPropFemaleVsMale`. We started the object name with the word `Model` to indicate that it stores the output of a statistical model. `ModelPropFemaleVsMale` is a complex object. However, when we print it with the `print()` command, we extract important information from the model object:

```{r RAndRStudio-Exerc100-701, exercise=FALSE, eval=TRUE}
print(ModelPropFemaleVsMale)
```

Reading the output above backward, starting with the last line, allows us to develop the final result step by step. The last line from the `ModelPropFemaleVsMale` object shows the different survival rates for female (`Prop1`) and male (`Prop2`) passengers. The difference between these two survival rates is about 0.55. The 95%-confidence interval for the difference ranges from 0.49 to 0.61. The probability of this difference being zero (both survival rates would be the same) is expressed by the P-value in line five of the output. As the P-value shows, the probability for the female and the male survival rate to be the same is almost zero.

We can summarize:

**The assertion from the Telegraph that chivalry was dead in 1912 during the Titanic disaster is most likely not correct.**

Now, you should be ready to pursue another small project with minimal guidance (no `Hints` in the code block below). How about you analyze the **different survival rates depending on if passengers traveled in first, second, or third class?**

First, create, based on the data frame `DataAnalysis`, three data frames containing only passengers in classes 1, 2, and 3, respectively. The object names followed by `...` are already provided in the *R* code block below.

Next, substitute the `...` in lines 5 -- 7 with the appropriate *R* code to calculate the survival rates for each of the three travel classes (note that the data frame `DataAnalysis` has already been loaded in the background):

```{r RAndRStudio-Exerc100-800, exercise=TRUE, exercise.setup="RAndRStudio-Exerc100-DataChivalryLoad", eval=FALSE, exercise.lines=15}
DataClass1= ...
DataClass2= ...
DataClass3= ...
  
Class1SurvRate= ...
Class2SurvRate= ...
Class3SurvRate= ...  

cat("Survival rate in class 1:", Class1SurvRate)
cat("Survival rate in class 2:", Class2SurvRate)
cat("Survival rate in class 3:", Class3SurvRate)
```

**Happy analysis!**

</clex>